---
title: "Physical Measures"
author: "jdt"
date: "2023-09-20"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background
Larner (1996) measured the weight and various physical measurements for 22 males aged 16–30. 
\pause \
Subjects were randomly chosen volunteers and all were in reasonably good health.  \
\pause
Subjects were requested to tense each muscle being measured slightly to ensure measurement consistency. \
\pause
The question of interest for these data is how weight can best be predicted from the other measurements.

## Dataset
```{r dataset}
library(readxl)
PhysicalMeasures <- read_xlsx(file.path("physicalmeasures.xlsx"))
summary(PhysicalMeasures)
```

## Scatterplot Matrix
```{r scatterplot matrix code, message=FALSE, warning=FALSE, fig.show='hide'}
library(GGally)
ggpairs(PhysicalMeasures, axisLabels = "none", 
        title = "Scatterplot Matrix of Physical Measures")
# corr codes
```

## Scatterplot Matrix
```{r scatterplot matrix, echo = FALSE}
ggpairs(PhysicalMeasures, axisLabels = "none", 
        title = "Scatterplot Matrix of Physical Measures")
```

## Linear Model of Mass with Forearm as the Predictor
```{r LM with all predictors code, message=FALSE, warning=FALSE, results=FALSE}
#waist=Physicalmeasures$Waist

fit_fore <- lm(Mass ~ Fore, data = PhysicalMeasures)
summary(fit_fore) # does not have VIFs included like SAS
plot(Mass ~ Fore, data = PhysicalMeasures)
abline(fit_fore)
```

```{r LM with all predictors, echo=FALSE}
## Linear Model of Mass with All Predictors

#fit_all <- lm(Mass ~ ., data = PhysicalMeasures)
#summary(fit_all) # does not have VIFs included like SAS
```

## Checking Regression Assumptions
Linear regression makes several assumptions about the data, such as :
\pause \
1) Linearity of the data: The relationship between the predictor (x) and the outcome (y) is assumed to be linear. 
\pause \
2) Normality of residuals: The residual errors are assumed to be normally distributed. 
\pause \
3) Homogeneity of residuals variance: The residuals are assumed to have a constant variance (homoscedasticity). 
\pause \
4) Independence of residuals error terms. 
\pause \
You should check whether or not these assumptions hold true. 

## Fit Diagnostics -- Predicted Means
```{r predicted values}
fitted(fit_fore) # predicted values
```
## Fit Diagnostics -- Residuals
```{r residuals}
residuals(fit_fore) # residuals
```
## Fit Diagnostics -- ANOVA Table
```{r anova}
summary(fit_fore) 
summary.aov(fit_fore)
```
All these assumptions and potential problems can be checked by producing some diagnostic plots visualizing the residual errors.

```{r}
plot(fit_fore)
```

## Potential Problems
1) Non-linearity of the outcome - predictor relationships
\pause \
2) Heteroscedasticity: Non-constant variance of error terms.
\pause  \
3) Presence of influential values in the data that can be:
\pause \
-Outliers: extreme values in the outcome (y) variable \
-High-leverage points: extreme values in the predictors (x) variable
\pause \

\hl

## Linear Regression with Waist as the Predictor
```{r}
fit_waist <- lm(Mass ~ Waist, data = PhysicalMeasures)
summary(fit_waist) # does not have VIFs included like SAS
plot(Waist ~ Fore, data = PhysicalMeasures)
abline(fit_waist)
fitted(fit_waist) # predicted values
residuals(fit_waist) # residuals
summary(fit_waist) 
summary.aov(fit_waist)
plot(fit_waist)
```