---
title: "Archosaurs: Linear Regression Example"
author: "Katie, Rita, and Chang"
date: "10/6/23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/katieclewett/Desktop")
brain <- read.csv('archosaur.csv', header = TRUE)
```

# Distribution of body weight
```{r}
library("ggplot2")
ggplot(brain, aes(x = Body.Weight)) + geom_histogram(bins = 16, color = "black", fill = "violet") + ggtitle("Histogram of Body Weight") + scale_x_continuous(limits = c(-100,90000))

ggplot(brain, aes(x = Body.Weight)) + geom_boxplot() + ggtitle("Boxplot of Body Weight")

xbar <- mean(brain$Body.Weight)
xbar
sd <- sd(brain$Body.Weight)
sd
n = 21
standard_error_mean <- sd/sqrt(n)
standard_error_mean
margin <- qt(0.975,df=n-1)*sd/sqrt(n)
lowerinterval <- xbar - margin
lowerinterval
upperinterval <- xbar + margin
upperinterval
```

# Distribution of brain weight
```{r}
library("ggplot2")
ggplot(brain, aes(x = Brain.Weight)) + geom_histogram(bins=10, color = "black", fill = "violet") + ggtitle("Histogram of Brain Weight") + scale_x_continuous(limits = c(0,400))

ggplot(brain, aes(x = Brain.Weight)) + geom_boxplot() + ggtitle("Boxplot of Brain Weight")

xbar <- mean(brain$Brain.Weight)
xbar
sd <- sd(brain$Brain.Weight)
sd
n = 21
standard_error_mean <- sd/sqrt(n)
standard_error_mean
margin <- qt(0.975,df=n-1)*sd/sqrt(n)
lowerinterval <- xbar - margin
lowerinterval
upperinterval <- xbar + margin
upperinterval
```

# Regression with body weight and brain weight prior to transformation
```{r}
plot(brain$Brain.Weight~brain$Body.Weight)
abline(lm(brain$Brain.Weight~brain$Body.Weight))
```

# Transform body weight and brain weight
```{r}
brain$log_body.wt = log(brain$Body.Weight)
brain$log_brain.wt = log(brain$Brain.Weight)
```

# Distribution of transformed body weight
```{r}
library("ggplot2")
ggplot(brain, aes(x = log_body.wt)) + geom_histogram(bins = 16, color = "black", fill = "violet") + ggtitle("Histogram of Transformed Body Weight") 

ggplot(brain, aes(x = log_body.wt)) + geom_boxplot() + ggtitle("Boxplot of Transformed Body Weight")

xbar <- mean(brain$log_body.wt)
xbar
sd <- sd(brain$log_body.wt)
sd
n = 21
standard_error_mean <- sd/sqrt(n)
standard_error_mean
margin <- qt(0.975,df=n-1)*sd/sqrt(n)
lowerinterval <- xbar - margin
lowerinterval
upperinterval <- xbar + margin
upperinterval
```

# Distribution of transformed brain weight
```{r}
library("ggplot2")
ggplot(brain, aes(x = log_brain.wt)) + geom_histogram(bins=10, color = "black", fill = "violet") + ggtitle("Histogram of Transformed Brain Weight") 

ggplot(brain, aes(x = log_brain.wt)) + geom_boxplot() + ggtitle("Boxplot of Transformed Brain Weight")

xbar <- mean(brain$log_brain.wt)
xbar
sd <- sd(brain$log_brain.wt)
sd
n = 21
standard_error_mean <- sd/sqrt(n)
standard_error_mean
margin <- qt(0.975,df=n-1)*sd/sqrt(n)
lowerinterval <- xbar - margin
lowerinterval
upperinterval <- xbar + margin
upperinterval
```

# Regression plot of transformed data
```{r}
plot(log_brain.wt~log_body.wt, dat = brain)
abline(lm(log_brain.wt~log_body.wt, dat = brain))
result<-lm(log_brain.wt~log_body.wt, dat = brain)
plot(result)
```

# Lack of fit test
```{r}
library("olsrr")
fit <- lm(log_brain.wt~log_body.wt, dat = brain)
ols_pure_error_anova(fit)
```

# Parameter estimates
```{r}
result<-lm(log_brain.wt~log_body.wt, dat = brain)
summary(result)
confint(result)
```

# Predict brain weight if body weight is 20
```{r}
x <- data.frame(20, log(20))
names(x) <- c("Body.Weight", "log_body.wt")
pred <- predict(result, x)
pred
exp(pred)
```

# Residual Plot
```{r}
result<-lm(log_brain.wt~log_body.wt, dat = brain)
res <- resid(result)
plot(fitted(result), res)
abline(0,0)
```