---
title: "Red Wine - linear"
author: "Katie, Rita, and Chang"
date: "2023-10-23"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/katieclewett/Desktop")
red <- read.csv("wineQualityReds.csv")
```

# Scatterplot Matrix
```{r scatterplot matrix code, message=FALSE, warning=FALSE, fig.show='hide'}
library("GGally")
ggpairs(red, axisLabels = "none", 
        title = "Scatterplot Matrix of Red Wines")
# corr codes
```

# Scatterplot Matrix
```{r scatterplot matrix, echo = FALSE}
ggpairs(red, axisLabels = "none", 
        title = "Scatterplot Matrix of Red Wines")
```

# Stepwise Regression
```{r}
library("MASS")
full.red <- lm(quality ~ . - id, data = red)
step.red <- stepAIC(full.red, direction = "both", trace = FALSE)
summary(step.red)
```

# Stepwise Regression Model
```{r}
library("car")
step <- lm(quality ~ vol_acidity + chlorides + free_sulfur + total_sulfur + pH + sulphates + alcohol, data = red)
summary(step)

# Check model for multicollinearity 
vif(step)
```

# Forward Selection
```{r}
library("MASS")
library("olsrr")
full.red <- lm(quality ~ . - id, data = red)
ols_step_forward_aic(full.red, details = TRUE)
# results in same model as stepwise regression
```

# Backward Elimination
```{r}
library("MASS")
library("olsrr")
full.red <- lm(quality ~ . - id, data = red)
ols_step_backward_aic(full.red, details = TRUE)
# results in same model as stepwise regression
```

# Standardize the variables
```{r}
library("tidyverse")
library("broom")
library("mosaic")

red_standardized <- 
  red %>%
  mutate(fix_acidity = scale(fix_acidity), vol_acidity = scale(vol_acidity), citric_acid = scale(citric_acid), sugar = scale(sugar), chlorides = scale(chlorides), free_sulfur = scale(free_sulfur), total_sulfur = scale(total_sulfur), density = scale(density), pH = scale(pH), sulphates = scale(sulphates), alcohol = scale(alcohol))
```

# Lasso Regression
```{r}
library("glmnet")
x = data.matrix(red_standardized [,c('fix_acidity', 'vol_acidity', 'citric_acid', 'sugar', 'chlorides', 'free_sulfur', 'total_sulfur', 'density', 'pH', 'sulphates', 'alcohol')])
y = red_standardized$quality
model <- glmnet(x, y, alpha = 1)
summary(model)
cv_model <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model$lambda.min
best_lambda
plot(cv_model)
best_model <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model)
```

# Lasso Regression Model
```{r}
library("car")
lassomod <- lm(quality ~ vol_acidity + sugar + chlorides + free_sulfur + total_sulfur + pH + sulphates + alcohol, data = red_standardized)
summary(lassomod)

# Check model for multicollinearity 
vif(lassomod)
```

# Lasso Regression max of 5 variables
```{r}
library("glmnet")
x = data.matrix(red_standardized [,c('fix_acidity', 'vol_acidity', 'citric_acid', 'sugar', 'chlorides', 'free_sulfur', 'total_sulfur', 'density', 'pH', 'sulphates', 'alcohol')])
y = red_standardized$quality
model <- glmnet(x, y, alpha = 1)
summary(model)
cv_model <- cv.glmnet(x, y, alpha = 1, dfmax = 4)
best_lambda <- cv_model$lambda.min
best_lambda
plot(cv_model)
best_model <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model)
```

# Lasso Regression Model max of 5 variables
```{r}
library("car")
lassomod5 <- lm(quality ~ vol_acidity + chlorides +  total_sulfur + sulphates + alcohol, data = red_standardized)
summary(lassomod5)

# Check model for multicollinearity 
vif(lassomod5)
```

# Ridge Regression
```{r}
library("glmnet")
x = data.matrix(red_standardized [,c('fix_acidity', 'vol_acidity', 'citric_acid', 'sugar', 'chlorides', 'free_sulfur', 'total_sulfur', 'density', 'pH', 'sulphates', 'alcohol')])
y = red_standardized$quality
model <- glmnet(x, y, alpha = 0)
summary(model)
cv_model <- cv.glmnet(x, y, alpha = 0)
best_lambda <- cv_model$lambda.min
best_lambda
plot(cv_model)
best_model <- glmnet(x, y, alpha = 0, lambda = best_lambda)
coef(best_model)
```

# Ridge Regression Model
```{r}
library("car")
ridgemod <- lm(quality ~ vol_acidity + chlorides + total_sulfur + pH + sulphates + alcohol, data = red_standardized)
summary(ridgemod)

# Check model for multicollinearity 
vif(ridgemod)
```

# Select Best Model with AIC
```{r}
library(AICcmodavg)
models <- list(step, lassomod, ridgemod, lassomod5)
mod.names <- c('Stepwise', 'Lasso', 'Ridge', 'Limited Lasso')
aictab(cand.set = models, modnames = mod.names)
# step is best model
```

# Select Best Model with BIC
```{r}
library("flexmix")
BIC(step)
BIC(lassomod)
BIC(ridgemod)
BIC(lassomod5)
# ridge is best model
# however, since the step model performed better with the adjusted r-squared and aic tests, we will move forward with that model
```

# From this point forward, we are using the stepwise regression model.

# Diagnostics
```{r}
library("broom")
diagnostics <- augment(step)
```

# Distribution of quality
```{r}
library("ggplot2")
ggplot(red, aes(x = quality)) + geom_histogram(bins = 10, color = "black", fill = "violet") + ggtitle("Histogram of Red Wine Quality") + scale_x_continuous(limits = c(0,10))

ggplot(red, aes(x = quality)) + geom_boxplot() + ggtitle("Boxplot of Red Wine Quality")

xbar <- mean(red$quality)
xbar
sd <- sd(red$quality)
sd
n = 1599
standard_error_mean <- sd/sqrt(n)
standard_error_mean
margin <- qt(0.975,df=n-1)*sd/sqrt(n)
lowerinterval <- xbar - margin
lowerinterval
upperinterval <- xbar + margin
upperinterval
```

# Plot Residuals
```{r}
library("ggplot2")
model_resid = step$residuals
ggplot(red, aes(x = model_resid)) + geom_histogram(bins = 15, color = "black", fill = "violet") + ggtitle("Histogram of Residuals") + scale_x_continuous(limits = c(-3,3))

ggplot(red, aes(x = model_resid)) + geom_boxplot() + ggtitle("Boxplot of Residuals")
```

# QQ Plot of Residuals - check normality
```{r}
qqnorm(model_resid)
qqline(model_resid)
```

# Multicollinearity Plot - check multicollinearity
```{r}
library("ggcorrplot")
red1 <- data.frame(red$vol_acidity, red$chlorides, red$free_sulfur, red$total_sulfur, red$pH, red$sulphates, red$alcohol)
corr_matrix = round(cor(red1), 2)
ggcorrplot(corr_matrix, hc.order = TRUE, type = "lower", lab = TRUE)
```

# Other Plots
```{r}
plot(quality ~ vol_acidity + chlorides + free_sulfur + total_sulfur + pH + sulphates + alcohol, data = red)

residual.quality = residuals(step)
predicted.quality = predict(step)
plot(residual.quality, predicted.quality)

plot(red$quality, predicted.quality)
reg = lm(predicted.quality ~ red$quality)
abline(reg)

plot(step, which = c(1,2,3,4,5,6))
```

# Residuals by vol_acidity
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = red$vol_acidity, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by chlorides
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = red$chlorides, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
``` 

# Residuals by free_sulfur
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = red$free_sulfur, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
``` 

# Residuals by total_sulfur
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = red$total_sulfur, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
``` 

# Residuals by pH
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = red$pH, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
``` 

# Residuals by sulphates
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = red$sulphates, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
``` 

# Residuals by alcohol
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = red$alcohol, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
``` 