---
title: "Physical Measures Project"
author: "Randy Nguyen, Emmie Jenkins, Sonish Lamsal"
date: "11/19/2020"
#papersize: a5
#fontsize: 8pt
output: html_document
#theme: "Berkeley"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background
Larner (1996) measured the weight and various physical measurements for 22 males aged 16–30. 
\pause \
Subjects were randomly chosen volunteers and all were in reasonably good health.  \
\pause
Subjects were requested to tense each muscle being measured slightly to ensure measurement consistency. \
\pause
The question of interest for these data is how weight can best be predicted from the other measurements.

## Dataset
```{r dataset}
library(readxl)
PhysicalMeasures <- read_xlsx(file.path("physicalmeasures.xlsx"))
```

## Scatterplot Matrix
```{r scatterplot matrix code, message=FALSE, warning=FALSE, fig.show='hide'}
library(GGally)
ggpairs(PhysicalMeasures, axisLabels = "none", 
        title = "Scatterplot Matrix of Physical Measures")
# corr codes
```

## Scatterplot Matrix
```{r scatterplot matrix, echo = FALSE}
ggpairs(PhysicalMeasures, axisLabels = "none", 
        title = "Scatterplot Matrix of Physical Measures")
```

## Linear Model of Mass with All Predictors
```{r LM with all predictors code, message=FALSE, warning=FALSE, results=FALSE}
fit <- lm(Mass ~ ., data = PhysicalMeasures)
summary(fit) # does not have VIFs included like SAS
```

## Linear Model of Mass with All Predictors
```{r LM with all predictors, echo=FALSE}
fit <- lm(Mass ~ ., data = PhysicalMeasures)
summary(fit) # does not have VIFs included like SAS
```

## Checking Regression Assumptions
Linear regression makes several assumptions about the data, such as :
\pause \
1) Linearity of the data: The relationship between the predictor (x) and the outcome (y) is assumed to be linear. \
\pause
2) Normality of residuals: The residual errors are assumed to be normally distributed. \
\pause
3) Homogeneity of residuals variance: The residuals are assumed to have a constant variance (homoscedasticity). \
\pause
4) Independence of residuals error terms. \
\pause
You should check whether or not these assumptions hold true. 

## Potential Problems
1) Non-linearity of the outcome - predictor relationships
\pause
2) Heteroscedasticity: Non-constant variance of error terms.
\pause
3) Presence of influential values in the data that can be:
\pause \
-Outliers: extreme values in the outcome (y) variable \
-High-leverage points: extreme values in the predictors (x) variable
\pause \
All these assumptions and potential problems can be checked by producing some diagnostic plots visualizing the residual errors.

## Fit Diagnostics
```{r predicted values}
fitted(fit) # predicted values
```

## Fit Diagnostics
```{r residuals}
residuals(fit) # residuals
```
## Fit Diagnostics
```{r fit diagnostics code, message=FALSE, warning=FALSE}
library(broom)
options(width = 80)
diagnostics <- augment(fit)
diagnostics[, 12:16]
```

## Fit Diagnostics
```{r anova}
anova(fit) 
```

## Fit Diagnostics
```{r diagnostic plot 1}
plot(fit, which = 1)
```

## Fit Diagnostics
```{r diagnostic plot 2}
plot(fit, which = 2)
```

## Fit Diagnostics
```{r diagnostic plot 3}
plot(fit, which = 3)
```

## Fit Diagnostics
```{r diagnostic plot 4}
plot(fit, which = 4)
```

## Fit Diagnostics
```{r diagnostic plot 5}
plot(fit, which = 5)
```

## Fit Diagnostics
```{r diagnostic plot 6}
plot(fit, which = 6)
```

## Residual by Regressors for Mass
```{r fit diagnostics code 1, fig.show='hide'}
ggplot(diagnostics) +
  geom_point(aes(x = Fore, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics 1, echo=FALSE}
ggplot(diagnostics) +
  geom_point(aes(x = Fore, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics code 2, fig.show='hide'}
ggplot(diagnostics) +
  geom_point(aes(x = Bicep, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics 2, echo=FALSE}
ggplot(diagnostics) +
  geom_point(aes(x = Bicep, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics code 3, fig.show='hide'}
ggplot(diagnostics) +
  geom_point(aes(x = Chest, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics 3, echo=FALSE}
ggplot(diagnostics) +
  geom_point(aes(x = Chest, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics code 4, fig.show='hide'}
ggplot(diagnostics) +
  geom_point(aes(x = Neck, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics 4, echo=FALSE}
ggplot(diagnostics) +
  geom_point(aes(x = Neck, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics code 5, fig.show='hide'}
ggplot(diagnostics) +
  geom_point(aes(x = Shoulder, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics 5, echo=FALSE}
ggplot(diagnostics) +
  geom_point(aes(x = Shoulder, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics code 6, fig.show='hide'}
ggplot(diagnostics) +
  geom_point(aes(x = Waist, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics 6, echo=FALSE}
ggplot(diagnostics) +
  geom_point(aes(x = Waist, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics code 7, fig.show='hide'}
ggplot(diagnostics) +
  geom_point(aes(x = Height, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics 7, echo=FALSE}
ggplot(diagnostics) +
  geom_point(aes(x = Height,y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics code 8, fig.show='hide'}
ggplot(diagnostics) +
  geom_point(aes(x = Calf, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics 8, echo=FALSE}
ggplot(diagnostics) +
  geom_point(aes(x = Calf, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics code 9, fig.show='hide'}
ggplot(diagnostics) +
  geom_point(aes(x = Thigh, y = .resid)) +
  geom_hline(yintercept=0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics 9, echo=FALSE}
ggplot(diagnostics) +
  geom_point(aes(x = Thigh, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics code 10, fig.show='hide'}
ggplot(diagnostics) +
  geom_point(aes(x = Head, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

## Residual by Regressors for Mass
```{r fit diagnostics 10, echo=FALSE}
ggplot(diagnostics) +
  geom_point(aes(x = Head, y = .resid)) +
  geom_hline(yintercept=0) +
  ylab("Residual") +
  theme_minimal()
```

## Evaluate Collinearity
```{r collinearity code, results=FALSE}
library(car)
options(width = 60)
vif(fit) # variance inflation factors
sqrt(vif(fit)) > 2 # Ice Cream Flavor
```

## Evaluate Collinearity
```{r collinearity, echo=FALSE}
vif(fit)
sqrt(vif(fit)) > 2 
```

## Test for Autocorrelated Errors
```{r autocorrelation}
durbinWatsonTest(fit)
```

## Global Test of Model Assumptions
```{r testing model assumptions code, warning=FALSE, results=FALSE}
#library(gvlma)
#gvmodel <- gvlma(fit)
#display.gvlmatests(gvmodel)
```

## Global Test of Model Assumptions
```{r testing model assumptions, echo=FALSE}
#display.gvlmatests(gvmodel)
```

## Backward Stepwise Model Selection Based on AIC
```{r backward stepwise code, results=FALSE}
selectedModel <- step(lm(Mass ~ ., data = PhysicalMeasures))
summary(selectedModel)
```

## Backward Stepwise Model Selection Based on AIC
```{r backward stepwise, echo=FALSE}
summary(selectedModel)
```

## Check VIFs for Multicollinearity
```{r multicollinearity}
print(all_vifs <- car::vif(selectedModel))
```

## Recursively select models with VIF < 4
```{r backward stepwise with VIFs code, results=FALSE}
signif_all <- names(all_vifs)
while(any(all_vifs > 4)){
  # get the var with max vif
  var_with_max_vif <- names(which(all_vifs == max(all_vifs)))  
  # remove
  signif_all <- signif_all[!(signif_all) %in% var_with_max_vif]  
  # new formula
  myFormula <- as.formula(paste("Mass ~ ", 
                             paste (signif_all, collapse=" + "), sep=""))  
  # re-build model with new formula
  selectedModel <- lm(myFormula, data = PhysicalMeasures)  
  all_vifs <- car::vif(selectedModel)
}
summary(selectedModel)
```

## Recursively select models with VIF < 4
```{r backward stepwise with VIFs, echo=FALSE}
signif_all <- names(all_vifs)
while(any(all_vifs > 4)){
  # get the var with max vif
  var_with_max_vif <- names(which(all_vifs == max(all_vifs)))  
  # remove
  signif_all <- signif_all[!(signif_all) %in% var_with_max_vif]  
  # new formula
  myFormula <- as.formula(paste("Mass ~ ", 
                             paste (signif_all, collapse=" + "), sep=""))  
  # re-build model with new formula
  selectedModel <- lm(myFormula, data = PhysicalMeasures)  
  all_vifs <- car::vif(selectedModel)
}
summary(selectedModel)
```

## P-value
The p-value for Head is large, so not significant
```{r p-value}
car::vif(selectedModel)
```
May want to leave it in to support the model

## Recursively Remove Non-significant predictors with VIFs Criteria
```{r backward stepwise with VIFs and p-value code, results=FALSE}
all_vars <- names(selectedModel[[1]])[-1]  # names of all X variables
# Get the non-significant vars
summ <- summary(selectedModel)  # model summary
pvals <- summ[[4]][, 4]  # get all p values
# init variables that aren't statistically significant
not_significant <- character() 
not_significant <- names(which(pvals > 0.1))
# remove 'intercept'. Optional!
not_significant <- not_significant[!not_significant %in% "(Intercept)"] 
# If there are any non-significant variables, 
while(length(not_significant) > 0){
  all_vars <- all_vars[!all_vars %in% not_significant[1]]
  # new formula
  myFormula <- as.formula(paste("Mass ~ ", 
    paste (all_vars, collapse = " + "), sep = ""))
  # re-build model with new formula
  selectedModel <- lm(myFormula, data = PhysicalMeasures)  
  # Get the non-significant vars.
  summ <- summary(selectedModel)
  pvals <- summ[[4]][, 4]
  not_significant <- character()
  not_significant <- names(which(pvals > 0.1))
  not_significant <- not_significant[!not_significant %in% "(Intercept)"]
}
summary(selectedModel)
```

## Recursively Remove Non-significant predictors with VIFs Criteria
```{r backward stepwise with VIFs and p-value, echo=FALSE}
all_vars <- names(selectedModel[[1]])[-1]  # names of all X variables
# Get the non-significant vars
summ <- summary(selectedModel)  # model summary
pvals <- summ[[4]][, 4]  # get all p values
not_significant <- character()  # init variables that aren't statistically significant
not_significant <- names(which(pvals > 0.1))
not_significant <- not_significant[!not_significant %in% "(Intercept)"] # remove 'intercept'. Optional!

# If there are any non-significant variables, 
while(length(not_significant) > 0){
  all_vars <- all_vars[!all_vars %in% not_significant[1]]
  myFormula <- as.formula(paste("Mass ~ ", paste (all_vars, collapse=" + "), sep=""))  # new formula
  selectedModel <- lm(myFormula, data = PhysicalMeasures)  # re-build model with new formula
  
  # Get the non-significant vars.
  summ <- summary(selectedModel)
  pvals <- summ[[4]][, 4]
  not_significant <- character()
  not_significant <- names(which(pvals > 0.1))
  not_significant <- not_significant[!not_significant %in% "(Intercept)"]
}
summary(selectedModel)
```

## Forward Stepwise
```{r forward stepwise code, results=FALSE, warning=FALSE, message=FALSE}
library(MASS); library(caret)
res.lm <- lm(Mass ~., data = PhysicalMeasures)
step <- stepAIC(res.lm, direction = "forward", trace = FALSE)
step
# Train the model
train.control <- trainControl(method = "cv", number = 10)
step.model <- train(Mass ~., data = PhysicalMeasures,
                    method = "lmStepAIC", 
                    trControl = train.control,
                    trace = FALSE
)
# Model accuracy
step.model$results
# Final model coefficients
step.model$finalModel
# Summary of the model
summary(step.model$finalModel)
```

## Forward Stepwise
```{r forward stepwise, echo=FALSE}
# Summary of the mode
summary(step.model$finalModel)
```

## Lasso
```{r lasso code, results=FALSE, warning=FALSE, message=FALSE}
library(ISLR); library(glmnet); library(dplyr); library(tidyr)
X = PhysicalMeasures[-1] %>% as.matrix()
y = PhysicalMeasures[,1] %>% as.matrix()

#must specify alpha=1 for lasso
fit_lasso = cv.glmnet(X, y, alpha=1)
#we can obtain the lasso estimates
predict(fit_lasso, s ="lambda.min", type = "coefficients")
```

## Lasso
```{r lasso, echo=FALSE}
library(ISLR); library(glmnet); library(dplyr); library(tidyr)
X = PhysicalMeasures[-1] %>% as.matrix()
y = PhysicalMeasures[,1] %>% as.matrix()

#must specify alpha=1 for lasso
fit_lasso = cv.glmnet(X, y, alpha=1)
#we can obtain the lasso estimates
predict(fit_lasso, s ="lambda.min", type = "coefficients")
```

## Elastic Net
```{r elastic net code, results=FALSE, warning=FALSE, message=FALSE}
#must specify alpha=1 for lasso
fit_elnet <- cv.glmnet(X, y, alpha=.5,)
predict(fit_elnet, s ="lambda.min", type = "coefficients")
```

## Elastic Net
```{r elastic net, echo=FALSE}
#must specify alpha=1 for lasso
fit_elnet <- cv.glmnet(X, y, alpha=.5,)
predict(fit_elnet, s ="lambda.min", type = "coefficients")
```

## Chosen Model: Parsimonious
```{r chosen model}
summary(selectedModel)
```
