---
title: "Simulated ROC"
author: "jdt"
date: "12/13/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
Set seed for the simulation
```{r}

  # clear the environment and set seed
rm(list = ls())
set.seed(12345)
```
Function to generate normal data.
For dis=0 ("Control") and dis=1 ("Disease"). The separation between the two groups is controlled by one's choice of c0, c1, sd_e0, and sd_e1. n0 and n1 are the sample sizes for the groups.

```{r}
library(ggplot2)

gen_Norm_data = function(c0,sd_e0,n0,c1,sd_e1,n1){
  eps0 = rnorm(n0,0,sd_e0)
  eps1 = rnorm(n1,0,sd_e1)
  y0 = c0 + eps0
  y1 = c1 + eps1
  
  data.frame(y = c(y0, y1), 
             dis = as.factor(c(rep(0,n0), rep(1,n1))))
}
dat1 = gen_Norm_data(1.5,1.5,200,2.5,1.5,200)

#plot density functions
ggplot(dat1, aes(color = dis, y)) + geom_density()
```
Create discrete variables for Y
```{r}
y=dat1[,1]
summary(y)
case=dat1[,2]
```

Create binary table with cutoff y > 2.5
```{r}
high_y = y > 2.5
tex1 = table(case,high_y)
addmargins(tex1)
prop.out = prop.table(tex1,1)
specificity = prop.out[1,1]
sensitivity = prop.out[2,2]
prop.out
sensitivity
specificity
TPR = sensitivity
FPR = 1 - specificity
TPR
FPR
```

Create binary table with cutoff y > 3.0
```{r}
high_y = y > 3.0
tex1 = table(case,high_y)
addmargins(tex1)
prop.out = prop.table(tex1,1)
specificity = prop.out[1,1]
sensitivity = prop.out[2,2]
prop.out
sensitivity
specificity
TPR = sensitivity
FPR = 1 - specificity
TPR
FPR
```
Instead of creating a table at each cutoff point one can construct a ROC plot for a continuous variable Y. Which is constructing using the pairs (TPR, FPR) at each cutoff point (in R). Sometimes (as in SAS) this curve is smoothed.
```{r}
library(ROCR)
hist(y)

pred = prediction(y,case)
perf=performance(pred, "tpr", "fpr")
plot(perf)
```


```{r}
library(caTools)
colAUC(y, case, plotROC=TRUE)

```






