---
title: "Physical Measures"
author: "Red Team"
date: "2023-11-02"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
PhysicalMeasures <- read_xlsx("physicalmeasures.xlsx")
```


**Background:**

Larner (1996) measured the weight and various physical measurements for 22 males aged 16–30. 

Subjects were randomly chosen volunteers and all were in reasonably good health.

Subjects were requested to tense each muscle being measured slightly to ensure measurement consistency. 

The question of interest for these data is how weight can best be predicted from the other measurements.


**Dataset**
```{r dataset}
summary(PhysicalMeasures)
```

Scatterplot Matrix
```{r scatterplot matrix code, message=FALSE, warning=FALSE, fig.show='hide'}
library(GGally)
ggpairs(PhysicalMeasures, axisLabels = "none", 
        title = "Scatterplot Matrix of Physical Measures")
```

Scatterplot Matrix
```{r scatterplot matrix, echo = FALSE}
ggpairs(PhysicalMeasures, axisLabels = "none", 
        title = "Scatterplot Matrix of Physical Measures")
```

Linear Model of Mass with Forearm as the Predictor
```{r}
fit_fore <- lm(Mass ~ Fore, data = PhysicalMeasures)
summary(fit_fore) # does not have VIFs included like SAS
plot(Mass ~ Fore, data = PhysicalMeasures)
abline(fit_fore)
```

**Checking Regression Assumptions**
Linear regression makes several assumptions about the data, such as :
 1) Linearity of the data: The relationship between the predictor (x) and the outcome (y) is assumed to be linear. 

 2) Normality of residuals: The residual errors are assumed to be normally distributed. 

 3) Homogeneity of residuals variance: The residuals are assumed to have a constant variance (homoscedasticity). 

 4) Independence of residuals error terms. 

 You should check whether or not these assumptions hold true. 
 
Fit Diagnostics -- Predicted Means
```{r}
fitted(fit_fore) # predicted values
```


Fit Diagnostics -- Residuals
```{r}
residuals(fit_fore) # residuals
```


Fit Diagnostics -- ANOVA Table
```{r}
summary(fit_fore) 
summary.aov(fit_fore)
```

```{r}
plot(fit_fore)
```

**Potential Problems**
 1) Non-linearity of the outcome - predictor relationships

 2) Heteroscedasticity: Non-constant variance of error terms.

 3) Presence of influential values in the data that can be:

 Outliers: extreme values in the outcome (y) variable 
 High-leverage points: extreme values in the predictors (x) variable


Linear Model of Mass with Sholder as the Predictor
```{r}
fit_shold <- lm(Mass ~ Shoulder, data = PhysicalMeasures)
summary(fit_shold) # does not have VIFs included like SAS
plot(Mass ~ Shoulder, data = PhysicalMeasures)
abline(fit_shold)
```

Fit Diagnostics -- Predicted Means
```{r}
fitted(fit_shold) # predicted values
```


Fit Diagnostics -- Residuals
```{r}
residuals(fit_shold) # residuals
```


Fit Diagnostics -- ANOVA Table
```{r}
summary(fit_shold) 
summary.aov(fit_shold)
```

```{r}
plot(fit_shold)
```

**Potential Problems**
 1) Non-linearity of the outcome - predictor relationships

 2) Heteroscedasticity: Non-constant variance of error terms.

 3) Presence of influential values in the data that can be:

 Outliers: extreme values in the outcome (y) variable 
 High-leverage points: extreme values in the predictors (x) variable


Linear Model of Mass with Neck as the Predictor
```{r}
fit_neck <- lm(Mass ~ Neck, data = PhysicalMeasures)
summary(fit_neck) # does not have VIFs included like SAS
plot(Mass ~ Neck, data = PhysicalMeasures)
abline(fit_neck)
```

 Fit Diagnostics -- Predicted Means
```{r}
fitted(fit_neck) # predicted values
```
 Fit Diagnostics -- Residuals
```{r}
residuals(fit_neck) # residuals
```
 Fit Diagnostics -- ANOVA Table
```{r}
summary(fit_neck) 
summary.aov(fit_neck)
```

```{r}
plot(fit_neck)
```

**Potential Problems**
 1) Non-linearity of the outcome - predictor relationships

 2) Heteroscedasticity: Non-constant variance of error terms.

 3) Presence of influential values in the data that can be:

 Outliers: extreme values in the outcome (y) variable 
 High-leverage points: extreme values in the predictors (x) variable


Linear Model of Mass with Calf as the Predictor
```{r}
fit_calf <- lm(Mass ~ Calf, data = PhysicalMeasures)
summary(fit_calf) 
plot(Mass ~ Calf, data = PhysicalMeasures)
abline(fit_calf)
```

Fit Diagnostics -- Predicted Means
```{r}
fitted(fit_calf) # predicted values
```


Fit Diagnostics -- Residuals
```{r}
residuals(fit_calf) # residuals
```


Fit Diagnostics -- ANOVA Table
```{r}
summary(fit_calf) 
summary.aov(fit_calf)
```

```{r}
plot(fit_calf)
```

**Potential Problems**
 1) Non-linearity of the outcome - predictor relationships

 2) Heteroscedasticity: Non-constant variance of error terms.

 3) Presence of influential values in the data that can be:

 Outliers: extreme values in the outcome (y) variable 
 High-leverage points: extreme values in the predictors (x) variable


Linear Model of Mass with Head as the Predictor
```{r}
fit_head <- lm(Mass ~ Head, data = PhysicalMeasures)
summary(fit_head) # does not have VIFs included like SAS
plot(Mass ~ Head, data = PhysicalMeasures)
abline(fit_head)
```

Fit Diagnostics -- Predicted Means
```{r}
fitted(fit_head) # predicted values
```


Fit Diagnostics -- Residuals
```{r}
residuals(fit_head) # residuals
```


Fit Diagnostics -- ANOVA Table
```{r}
summary(fit_head) 
summary.aov(fit_head)
```

```{r}
plot(fit_head)
```


**Potential Problems**
 1) Non-linearity of the outcome - predictor relationships

 2) Heteroscedasticity: Non-constant variance of error terms.

 3) Presence of influential values in the data that can be:

 Outliers: extreme values in the outcome (y) variable 
 High-leverage points: extreme values in the predictors (x) variable

