---
title: "Physical Measure for Selection and Diagnostic"
author: "Red Team"
---

# Loading Libraries & Data Set
```{r}
library(GGally)
library(readxl)
library("MASS")
library("broom")
library("glmnet")

dat <- read_xlsx("physicalmeasures.xlsx")
```
# Scatterplot Matrix for Physical Measures
```{r}
ggpairs(dat, axisLabels = "none", 
        title = "Scatterplot Matrix of Physical Measures")
```

# Multiple Regression Model
```{r}
step <- lm(formula = Mass ~ Fore + Bicep + Chest + Neck + Shoulder +
                   Waist + Height + Calf + Thigh + Head , data = dat)
summary(step)

info <- augment(step)
```

# Residuals for Fore
```{r}
ggplot(info) +
  geom_point(aes(x = dat$Fore, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals for Bicep
```{r}
ggplot(info) +
  geom_point(aes(x = dat$Bicep, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals for Chest
```{r}
ggplot(info) +
  geom_point(aes(x = dat$Chest, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals for Neck
```{r}
ggplot(info) +
  geom_point(aes(x = dat$Neck, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals for Shoulder
```{r}
ggplot(info) +
  geom_point(aes(x = dat$Shoulder, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals for Waist
```{r}
ggplot(info) +
  geom_point(aes(x = dat$Waist, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals for Height
```{r}
ggplot(info) +
  geom_point(aes(x = dat$Height, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals for Calf
```{r}
ggplot(info) +
  geom_point(aes(x = dat$Calf, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals for Thigh
```{r}
ggplot(info) +
  geom_point(aes(x = dat$Thigh, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals for Head
```{r}
ggplot(info) +
  geom_point(aes(x = dat$Head, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Stepwise Regression Model
```{r}
model <- lm(Mass ~ ., data = dat)
step <- stepAIC(model, direction = "both", 
                      trace = FALSE)
summary(step)
```


```{r}
mass_res = residuals(step)
mass_pred= predict(step)
plot(mass_res, mass_pred)

plot(dat$Mass, mass_pred)
reg = lm(mass_pred ~ dat$Mass)
abline(reg)

plot(step, which = c(1,2,3,4,5,6))
```

# Lasso Regression Model
```{r}
body = data.matrix(dat [,c('Fore', 'Bicep', 'Chest', 'Neck', 'Shoulder',
                                     'Waist','Height', 'Calf', 'Thigh', 'Head')])
weight = dat$Mass
model_lasso <- glmnet(body, weight, alpha = 0)
summary(model_lasso)

model_two <- cv.glmnet(body, weight, alpha = 0)
top_lambda <- model_two$lambda.min
top_lambda
plot(model_two)

top <- glmnet(body, weight, alpha = 0, lambda = top_lambda)
coef(top)
```

# Ridge Regression Model
```{r}
body = data.matrix(dat [,c('Fore', 'Bicep', 'Chest', 'Neck', 'Shoulder',
                                     'Waist','Height', 'Calf', 'Thigh', 'Head')])
weight = dat$Mass
model_ridge <- glmnet(body, weight, alpha = 1)
summary(model_ridge)

model_two <- cv.glmnet(body, weight, alpha = 1)
top_lambda <- model_two$lambda.min
top_lambda
plot(model_two)

top <- glmnet(body, weight, alpha = 1, lambda = top_lambda)
coef(top)
```

