---
title: "SA Heart Study Data"
author: "author"
date: "date"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("repos" = c(CRAN = "https://cran.rstudio.com"))

```

Needed Packages
```{r, message=FALSE}
if(!require(knitr)){install.packages("knitr")}
if(!require(dplyr)){install.packages("dplyr")}
if (!require("readr")) install.packages("readr", dep=FALSE)
if (!require("gridExtra")) install.packages("gridExtra", dep=TRUE)

if(!require(FSA)){install.packages("FSA")}
if(!require(ggplot2)){install.packages("ggplot2")}
if (!require("mosaic")) install.packages("mosaic", dep=FALSE)
if (!require("nortest")) install.packages("nortest", dep=TRUE)
if (!require("epitools")) install.packages("epitools", dep=TRUE)
if (!require("prettyR")) install.packages("prettyR", dep=TRUE)
if (!require("rms")) install.packages("rms", dep=TRUE)
# add other as needed
```
## Observational Cohort Data

```{r, message=FALSE}
heart = read.csv('SAHeart_Disease.csv', header = TRUE)
#                       fileEncoding = 'UTF-8-BOM') #latest version
heart = data.frame(heart)


library(tidyr)
#Remove rows with NA's using drop_na()
heart <- heart %>% drop_na()

summary(heart)
dim(heart)
```

The objective is to model the probability that of CHD (CHD = '1'). 

Instead of creating a table at each cutoff point one can construct a ROC plot for a continuous variable Y. Which is constructing using the pairs (TPR, FPR) at each cutoff point (in R). Sometimes (as in SAS) this curve is smoothed.

Using ldl
```{r}
library(ROCR)
chd=heart$chd
adiposity = heart$adiposity
ldl = heart$ldl
hist(ldl)

pred = prediction(ldl,chd)

perf=performance(pred, "tpr", "fpr")
plot(perf)

library(caTools)

colAUC(ldl, chd, plotROC=TRUE)
```


Using adiposity
```{r}

hist(adiposity)

pred = prediction(adiposity,chd)

perf=performance(pred, "tpr", "fpr")
plot(perf)

library(caTools)

colAUC(adiposity, chd, plotROC=TRUE)
```

Using ldl and adiposity
```{r}

mod = ldl + adiposity
pred = prediction(mod,chd)

perf=performance(pred, "tpr", "fpr")
plot(perf)

library(caTools)

colAUC(mod, chd, plotROC=TRUE)
```
