---
title: "Red-Team-Archosaur"
author: "Red Team"
output: pdf_document
---

```{r}
# Loading the necessary packages & libraries
library("ggplot2")
library("olsrr")

# Loading the "archosaur" data set
data <- read.csv('archosaur.csv', header = TRUE)
```

# Body weight Graphs
```{r}
ggplot(data, aes(x = Body.Weight)) +
  geom_histogram(bins = 16, color = "blue", fill = "red") +
  ggtitle("Histogram of Body Weight") + scale_x_continuous(limits = c(0,90000))

ggplot(data, aes(x = Body.Weight)) +
  geom_boxplot() +
  ggtitle("Boxplot of Body Weight")
```

# Calculating the mean of Body weight
```{r}
xbar <- mean(data$Body.Weight)
xbar
```

# Calculating the standard deviation of body weight
```{r}
sd <- sd(data$Body.Weight)
sd
# Set n to a value of 21
n = 21
```

# Calculating the standard error mean of the body weight
```{r}
standard_error_mean <- sd/sqrt(n)
standard_error_mean

margin <- qt(0.975,df = n-1) * sd/sqrt(n)

lowerinterval <- xbar - margin
lowerinterval

upperinterval <- xbar + margin
upperinterval
```

# Brain Weight graphs
```{r}
ggplot(data, aes(x = Brain.Weight)) +
  geom_histogram(bins = 10, color = "blue", fill = "red") +
  ggtitle("Histogram of the Brain Weight") +
  scale_x_continuous(limits = c(0,400))

ggplot(data, aes(x = Brain.Weight)) +
  geom_boxplot() +
  ggtitle("Boxplot of the Brain Weight")
```

# Calculating the mean of the Brain Weight
```{r}
xbar <- mean(data$Brain.Weight)
xbar
```

# Calculating the standard deviation of the Brain Weight
```{r}
sd <- sd(data$Brain.Weight)
sd

# Set n to a value of 21
n = 21
```

# Calculating the standard error mean of the brain weight
```{r}
standard_error_mean <- sd/sqrt(n)
standard_error_mean

margin <- qt(0.975,df = n-1)*sd/sqrt(n)

lowerinterval <- xbar - margin
lowerinterval

upperinterval <- xbar + margin
upperinterval
```

# Regression Model between the body weight and data weight before the transformation occurs
```{r}
plot(data$Brain.Weight~data$Body.Weight)
abline(lm(data$Brain.Weight~data$Body.Weight))
```

# Transform ing the body weight and data weight
```{r}
data$log_body.wt = log(data$Body.Weight)
data$log_data.wt = log(data$Brain.Weight)
```

# Transformed body weight Graphs
```{r}
ggplot(data, aes(x = log_body.wt)) +
  geom_histogram(bins = 15, color = "blue", fill = "red") + 
  ggtitle("Histogram of Transformed Body Weight") 

ggplot(data, aes(x = log_body.wt)) + geom_boxplot() +
  ggtitle("Boxplot of Transformed Body Weight")
```

# Calculating the mean for the transformed body weight
```{r}
xbar <- mean(data$log_body.wt)
xbar
```

# Calculating the standard deviation for the transformed body weight
```{r}
sd <- sd(data$log_body.wt)
sd

# Setting the n to a value of 21
n = 21
```

# Calculating the standard error mean for the transformed body weight
```{r}
standard_error_mean <- sd/sqrt(n)
standard_error_mean

margin <- qt(0.975,df = n-1) * sd/sqrt(n)

lowerinterval <- xbar - margin
lowerinterval

upperinterval <- xbar + margin
upperinterval
```

# Transformed brain weight graphs
```{r}
ggplot(data, aes(x = log_data.wt)) +
  geom_histogram(bins = 10, color = "blue", fill = "red") +
  ggtitle("Histogram of Transformed Brain Weight") 

ggplot(data, aes(x = log_data.wt)) +
  geom_boxplot() +
  ggtitle("Boxplot of Transformed Brain Weight")
```

# Calculating the mean for the transformed brain weight
```{r}
xbar <- mean(data$log_data.wt)
xbar
```

# Calculating the standard deviation for the transformed brain weight
```{r}
sd <- sd(data$log_data.wt)
sd

# Setting n to a value of 21
n = 21
```

# Calculating the standard mean for the transformed brain weight
```{r}
standard_error_mean <- sd/sqrt(n)
standard_error_mean

margin <- qt(0.975,df = n-1)*sd/sqrt(n)

lowerinterval <- xbar - margin
lowerinterval

upperinterval <- xbar + margin
upperinterval
```

# Regression Model of the transformed data
```{r}
plot(log_data.wt~log_body.wt, dat = data)
abline(lm(log_data.wt~log_body.wt, dat = data))
result<-lm(log_data.wt~log_body.wt, dat = data)
plot(result)
```

# Creating a lack of fit test
```{r}
lack <- lm(log_data.wt~log_body.wt, dat = data)
ols_pure_error_anova(lack)
```


# Parameter estimates
```{r}
ans <- lm(log_data.wt~log_body.wt, dat = data)
summary(ans)
confint(ans)
```

# Estimating the brain weight if the body size is set to 20 kilograms
```{r}
value <- data.frame(20, log(20))
names(value) <- c("Body.Weight", "log_body.wt")
pred <- predict(ans, value)
pred
exp(pred)
```

# Residual Model Graohs
```{r}
ans <-lm(log_data.wt~log_body.wt, dat = data)
res <- resid(ans)
plot(fitted(ans), res)
abline(0,0)
```
