Loading in necessary data set
```{r}
library("ggplot2")
library("dplyr")
library("arsenal")
library("tidyverse")
library("rstatix")
library("ggpubr")
library("mosaic")
library("dgof")
```

Loading in the data set
```{r}
dat <- read.csv("baseball.csv", header=TRUE)
```

# Problem 1
```{r}
catcher <- filter(dat, Position == 'C')
CrHits2 <- dat$CrHits*dat$CrHits
in_fielder <- filter(dat, Position == '1B' | Position == '2B' |
                       Position == 'SS' | Position == '3B')
out_fielder <- filter(dat, Position == 'CF' | Position == 'RF' |
                        Position == 'LF' | Position == 'OF')
```

```{r}
leagueDiv <- table(dat$League, dat$Division)
div <- freqlist(leagueDiv)
summary(div)
```

```{r}
df <- filter(dat, League == 'American')
leagueDiv2 <- table(df$Division)
div2 <- freqlist(leagueDiv2)
summary(div2)
```

```{r}
df <- subset(df, select = c(Salary, Team, Division, YrMajor, logSalary, nAtBat,
                                nBB, nError, nHits, nHome, nRuns))
print(df[order(df$Division), ])
subDiv <- 
  df |>
  group_by(Division) |>
  summarize(c(total_count = n()), (mean_hits = mean(nHits)), (sd_hits = sd(nHits)),
            (se_hits = sd(nHits)/sqrt(n())), (min_hits = min(nHits)), (max_hits = max(nHits)))
print(subDiv)
```

```{r}
testSample <- t.test(df$nHits ~ df$Division, var.equal = TRUE)
print(testSample)
```

```{r}
t.test(df$nHits ~ df$Division)
```

Plot for West
```{r}
westDiv <- filter(df, Division == 'West')

qqnorm(westDiv$nHits)

ggplot(westDiv, aes(x = nHits)) +
  geom_boxplot()

ggplot((westDiv), aes(x = nHits)) +
  geom_histogram(aes(y = after_stat(density)), fill = 'green', col = 'black', bins = 20) +
  scale_x_continuous(limits = c(-50,300)) +
  stat_function(fun = dnorm, args = list(mean = mean(westDiv$nHits), sd = sd(westDiv$nHits)))
```

Plot of East
```{r}
eastDiv <- filter(df, Division == 'East')

qqnorm(eastDiv$nHits)

ggplot(eastDiv, aes(x = nHits)) +
  geom_boxplot()

ggplot((eastDiv), aes(x = nHits)) +
  geom_histogram(aes(y = after_stat(density)), fill = 'green', col = 'black', bins = 20) +
  scale_x_continuous(limits = c(-50,300)) +
  stat_function(fun = dnorm, args = list(mean = mean(eastDiv$nHits), sd=sd(eastDiv$nHits)))
```

```{r}
df |>
  group_by(Division) |>
  get_summary_stats(nHits, type = "median_iqr")

box <- ggboxplot(df, x = "Division", y = "nHits", ylab = "nHits", xlab = "Division")
box

wilcox.test(df$nHits~df$Division)

kruskal.test(df$nHits~df$Division)
```
Based on the results that we have calculated above, we found a small difference 
between the median and mean hits for each of the divisions. The p-values we have
calculated for the t-test shows that the results are not significant and concluding
that the mean and median have different values and results. 

# Problem 2
```{r}
df2 <- filter(dat, League == 'National' | Division == 'East')
dim(df2)
new_df = subset(df2, select = c(CrAtBat, CrBB, CrHits, CrHits2,CrRbi, CrRuns,
                                 Div, Salary, YrMajor, nBB, nHits, nHome)); 
```

```{r}
n_hits = df2$nHits
n_home = (df2$CrHits) / (df2$YrMajor)

summary(n_hits)
summary(n_home )

ggplot(df2, aes(x = n_hits, y = n_home)) +
  geom_boxplot()

t.test(n_hits,n_home)
wilcox.test(n_hits,n_home)
```
We changed the variable of interest by comparing the National League East with career
variables. Based on the results, we found that the results of the p-value from 
the t-test is significant, meaning that the there is no difference between
the measures of central tendency. 
