---
title: "Physical Measures Diagnostics"
author: "Katie, Rita, and Chang"
date: "2023-10-16"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/katieclewett/Desktop")
library(readxl)
PhysicalMeasures <- read_xlsx("physicalmeasures.xlsx")
```

# Scatterplot Matrix
```{r scatterplot matrix code, message=FALSE, warning=FALSE, fig.show='hide'}
library(GGally)
ggpairs(PhysicalMeasures, axisLabels = "none", 
        title = "Scatterplot Matrix of Physical Measures")
# corr codes
```

# Scatterplot Matrix
```{r scatterplot matrix, echo = FALSE}
ggpairs(PhysicalMeasures, axisLabels = "none", 
        title = "Scatterplot Matrix of Physical Measures")
```

# Stepwise Regression
```{r}
library("MASS")
full.phys <- lm(Mass ~ ., data = PhysicalMeasures)
step.phys <- stepAIC(full.phys, direction = "both", 
                      trace = FALSE)
summary(step.phys)
```


# Multiple regression with selected model (output of stepwise regression)
```{r}
stepmodel <- lm(formula = Mass ~ Fore + Waist + Height + Calf + Thigh + Head, data = PhysicalMeasures)
summary(stepmodel)
```

# Diagnostics
```{r}
library("broom")
diagnostics <- augment(stepmodel)
```

# Plots
```{r}
residual.mass = residuals(stepmodel)
predicted.mass = predict(stepmodel)
plot(residual.mass, predicted.mass)

plot(PhysicalMeasures$Mass, predicted.mass)
reg = lm(predicted.mass ~ PhysicalMeasures$Mass)
abline(reg)

plot(stepmodel, which = c(1,2,3,4,5,6))
```

# Residuals by Fore
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Fore, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by Waist
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Waist, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by Height
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Height, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by Calf
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Calf, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by Thigh
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Thigh, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by Head
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Head, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Ridge Regression
```{r}
library("glmnet")
x = data.matrix(PhysicalMeasures [,c('Fore', 'Bicep', 'Chest', 'Neck', 'Shoulder', 'Waist', 'Height', 'Calf', 'Thigh', 'Head')])
y = PhysicalMeasures$Mass
model <- glmnet(x, y, alpha = 0)
summary(model)
cv_model <- cv.glmnet(x, y, alpha = 0)
best_lambda <- cv_model$lambda.min
best_lambda
plot(cv_model)
best_model <- glmnet(x, y, alpha = 0, lambda = best_lambda)
coef(best_model)
```

# Lasso Regression
```{r}
library("glmnet")
x = data.matrix(PhysicalMeasures [,c('Fore', 'Bicep', 'Chest', 'Neck', 'Shoulder', 'Waist', 'Height', 'Calf', 'Thigh', 'Head')])
y = PhysicalMeasures$Mass
model <- glmnet(x, y, alpha = 1)
summary(model)
cv_model <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model$lambda.min
best_lambda
plot(cv_model)
best_model <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model)
```

# Multiple regression with selected model (output of stepwise regression)
```{r}
lassomodel <- lm(formula = Mass ~ . - Bicep - Neck, data = PhysicalMeasures)
summary(lassomodel)
```

# Diagnostics
```{r}
library("broom")
diagnostics <- augment(lassomodel)
```

# Plots
```{r}
residual.mass = residuals(lassomodel)
predicted.mass = predict(lassomodel)
plot(residual.mass, predicted.mass)

plot(PhysicalMeasures$Mass, predicted.mass)
reg = lm(predicted.mass ~ PhysicalMeasures$Mass)
abline(reg)

plot(lassomodel, which = c(1,2,3,4,5,6))
```

# Residuals by Fore
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Fore, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by Chest
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Chest, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by Shoulder
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Shoulder, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by Waist
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Waist, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by Height
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Height, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by Calf
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Calf, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by Thigh
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Thigh, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# Residuals by Head
```{r}
library("GGally")
ggplot(diagnostics) +
  geom_point(aes(x = PhysicalMeasures$Head, y = .resid)) +
  geom_hline(yintercept = 0) +
  ylab("Residual") +
  theme_minimal()
```

# The stepwise model is a better fit
# The stepwise model has a higher adjusted r-squared value