---
title: "Red Wines - upper"
author: "Katie, Rita, and Chang"
date: "2023-11-01"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/katieclewett/Desktop")
red <- read.csv("wineQualityReds.csv")
```

# Scatterplot Matrix
```{r scatterplot matrix code, message=FALSE, warning=FALSE, fig.show='hide'}
library("GGally")
ggpairs(red, axisLabels = "none", 
        title = "Scatterplot Matrix of Red Wines")
# corr codes
```

# Scatterplot Matrix
```{r scatterplot matrix, echo = FALSE}
ggpairs(red, axisLabels = "none", 
        title = "Scatterplot Matrix of Red Wines")
```

# Create Binary Dependent Variable
```{r}
red$highquality = factor((red$quality >= 6))
red$highquality <- as.integer(as.logical(red$highquality))
```

# Create Test and Training Data
```{r}
library("caTools")
set.seed = 100
split = sample.split(red$highquality, SplitRatio = 0.6)
train = subset(red, split == TRUE)
test = subset(red, split == FALSE)
print(dim(train)); print(dim(test))
```

# Descriptive Statistics
```{r}
library("Rmisc")

sum = summary(red)
sum
```

# Plot high quality vs low quality distribution
```{r}
hist (red$highquality)
```

# Random Forest 
```{r}
library("randomForest")
library("caret")
library("e1071")
library("rpart")

rf <- randomForest(highquality ~ . - quality, data = train, mtry = 4, importance = TRUE, ntree = 50, na.action = na.omit)
print(rf)
varImpPlot(rf)

# predictions on test set
set.seed(100)
predictTest = predict(rf, newdata = test, type = "response")

# confusion matrix on test set
table(test$highquality, predictTest >= 0.5)
520/nrow(test)

# the model is accurate 81.3 percent of the time
```

# Random Forest Model
```{r}
# Logit
randomforestmodlogit <- glm(highquality ~ alcohol + sulphates + total_sulfur + vol_acidity, data = red, family = "binomial"(link = "logit"))
summary(randomforestmodlogit)

# Cloglog
randomforestmodcloglog <- glm(highquality ~ alcohol + sulphates + total_sulfur + vol_acidity, data = red, family = "binomial"(link = "cloglog"))
summary(randomforestmodcloglog)

# The logit model performed better with a lower AIC value
```

# Cart
```{r}
library("caret")
library("e1071")
library("rpart")
library("rpart.plot")

cartmodel = rpart(highquality ~ . - quality, data = train)
print(cartmodel)
prp(cartmodel)

# predictions on test set
set.seed(100)
predictTest = predict(cartmodel, newdata = test, type = "matrix")

# confusion matrix on test set
table(test$highquality, predictTest >= 0.5)
```

# Cart Model
```{r}
# Logit
cartmodlogit <- glm(highquality ~ alcohol + sulphates + total_sulfur + fix_acidity, data = red, family = "binomial"(link = "logit"))
summary(cartmodlogit)

# Cloglog
cartmodcloglog <- glm(highquality ~ alcohol + sulphates + total_sulfur + fix_acidity, data = red, family = "binomial"(link = "cloglog"))
summary(cartmodcloglog)

# The logit model performed better with the lower AIC value
```

# Compare best logit model with AIC
```{r}
library("AICcmodavg")
models <- list(randomforestmodlogit, cartmodlogit)
mod.names <- c('RandomForest', 'Cart')
aictab(cand.set = models, modnames = mod.names)

# The random forest logit model performed the best
```

# Compare best model with BIC
```{r}
library("flexmix")
BIC(randomforestmodlogit)
BIC(randomforestmodcloglog)
BIC(cartmodlogit)
BIC(cartmodcloglog)

# The random forest logit model performed the best
```

# Confusion matrix for random forest logit model
```{r}
confusionred = predict(randomforestmodlogit, newdata = red, type = "response")

# confusion matrix on test set
table(red$highquality, confusionred >= 0.5)
```

# Predictions for random forest logit model
```{r}
pred_test <- predict(randomforestmodlogit, test, type = "response")

pred_test
```

# Model Diagnostics
```{r}
accuracy = (548+639)/(548+196+216+639)
accuracy

sensitivity = 639/(639+196)
sensitivity

specificity = 548/(548+216)
specificity
```

# AUC and ROC
```{r}
library("pROC")

test_prob = predict(randomforestmodlogit, test, type = "response")

test_roc = roc(test$highquality ~ test_prob, plot = TRUE, print.auc = TRUE)

as.numeric(test_roc$auc)
```

# AUC and ROC with just one variable
```{r}
library("ggplot2")

simple <- glm(highquality ~ vol_acidity, data = red, family = "binomial"(link = "logit"))
summary(simple)

ggplot(red, aes(x = vol_acidity, y = highquality)) +geom_point()+stat_smooth(method="glm", color="green", se=TRUE, method.args = list(family=binomial))

test_prop1 = predict(simple, red, type = "response")

test_roc1 = roc(red$highquality ~ test_prop1, plot = TRUE, print.auc = TRUE)

as.numeric(test_roc1$auc)
```