---
title: "Baseball Careers"
author: "author"
date: "date"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("repos" = c(CRAN = "https://cran.rstudio.com"))

```

# Load necessary libraries

```{r, message=FALSE}
if(!require(knitr)){install.packages("knitr")}
if(!require(dplyr)){install.packages("dplyr")}
if (!require("readr")) install.packages("readr", dep=FALSE)
if (!require("gridExtra")) install.packages("gridExtra", dep=TRUE)
if (!require("stats")) install.packages("stats", dep=TRUE)
if(!require(FSA)){install.packages("FSA")}
if(!require(ggplot2)){install.packages("ggplot2")}
if (!require("mosaic")) install.packages("mosaic", dep=FALSE)
if (!require("nortest")) install.packages("nortest", dep=TRUE)
if (!require("epitools")) install.packages("epitools", dep=TRUE)
if (!require("prettyR")) install.packages("prettyR", dep=TRUE)
if (!require("rms")) install.packages("rms", dep=TRUE)
# add others as needed
```
## Read the baseball data (assuming you have it in a CSV file)

```{r}
baseball <- read.csv("baseball.csv")
dim(baseball)
```

# Data for Problem 1 
 
Comparing American League East vs
American League West using 1986 data for seasonal hits


# Create a subset of the data for American League

```{r}
problem1 <- baseball %>%
  filter(League == "American") %>%
  select(Salary, Team, Division, YrMajor, logSalary, nAtBat, nBB, nError, nHits, nHome, nRBI, nRuns)

# Frequency table for Division
division_freq <- table(problem1$Division)
print(division_freq)
```

# Sort the data by Division
```{r}
problem1 <- problem1 %>%
  arrange(Division)

# Test for means assuming normal data (t-test)
t_test_result <- t.test(problem1$nHits ~ problem1$Division)
print(t_test_result)

# Test for means assuming non-normal data (Wilcoxon test)
wilcoxon_test_result <- wilcox.test(problem1$nHits ~ problem1$Division)
print(wilcoxon_test_result)

# Test for means assuming non-normal data (KS test)
ks_test_result <- ks.test(problem1$nHits ~ problem1$Division)
print(ks_test_result)
```

# Subset the data for problem 2 

Comparing present season
number of hits with past average number of hits for teams 
in National League East

```{r}
prob2 = subset(baseball, Div == "NE")
#summary(prob2)
dim(prob2)
#summary(prob2)
x=prob2$nHits                 #present number of hits
y = prob2$CrHits/prob2$YrMajor 
                              #past average number of hits

favstats(x)
favstats(y)

boxplot(x,y, data=prob2)
t.test(x,y)
wilcox.test(x,y)
ks.test(x,y)
```
