---
title: "Birth Weight"
author: "Team C"
date: "11/03/2021"
output: latex_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("repos" = c(CRAN = "https://cran.rstudio.com"))

```

Needed Packages
```{r, message=FALSE}

if(!require(FSA)){install.packages("FSA")}
if(!require(ggplot2)){install.packages("ggplot2")}
if (!require("mosaic")) install.packages("mosaic", dep=FALSE)
if (!require("nortest")) install.packages("nortest", dep=TRUE)
if (!require("epitools")) install.packages("epitools", dep=TRUE)
if (!require("prettyR")) install.packages("prettyR", dep=TRUE)
if (!require("rms")) install.packages("rms", dep=TRUE)
library(tidyverse)
# add other as needed
```

## Problem -- Infant Birth Weight Data

Read data into R
```{r}
# this data came from SASHELP.BWEIGHT
bw <- read.csv('bwgt.csv')
bw <- data.frame(bw)
#summary(bw)
bw = transform(bw, AgeGroup.f = as.factor(AgeGroup))
bw = transform(bw, Race.f = as.factor(Race))
bw = transform(bw, Drinking.f = as.factor(Drinking))
bw = transform(bw, Death.f = as.factor(Death))
bw = transform(bw, Smoking.f = as.factor(Smoking))
bw = transform(bw, SomeCollege.f = as.factor(SomeCollege))
bw = transform(bw, LowBirthWgt.f = as.factor(bw$LowBirthWgt))
```


\newpage
# Code 1 - Data Summary

```{r}
summary(bw)
summary(bw$LowBirthWgt)
summary(bw$Smoking)
```



```{r message=FALSE, warning=FALSE}
mytab <- tally(~ Race.f | AgeGroup.f, data=bw)
addmargins(mytab)

```


```{r message=FALSE, warning=FALSE}
mytab1 <- tally(~ Married | AgeGroup.f, data=bw)
addmargins(mytab1)

```


```{r  message=FALSE, warning=FALSE}
mytab2 <- tally(~ LowBirthWgt.f | Race.f, data=bw)
addmargins(mytab2)
```


\newpage

# Code 2

```{r}
bw <- read.csv('bwgt.csv')
# Create a new smaller data set of size 2500
set.seed(1568)
ssize <- sample(1:nrow(bw),round(nrow(bw) * .8092))
new_bwgt <- bw %>% slice(ssize)
new_bwgt <-new_bwgt  %>% na_if("") %>%  na.omit
 
# Make the following columns 
new_bwgt <- new_bwgt %>% mutate( AgeGroup.f = factor(AgeGroup), Race.f = factor(Race), 
Drinking.f = factor(Drinking), Death.f = factor(Death),Smoking.f = factor(Smoking),
SomeCollege.f = factor(SomeCollege), LowBirthWgt.f = factor(LowBirthWgt))
```

### Test for Association between Low Birth Weight and smoking
```{r}
#Risk ratio
RR1<-riskratio(x=new_bwgt$Smoking.f, y=new_bwgt$LowBirthWgt.f)
RR1$measure
RR1$p.value
```

Since 95% C.I. does contain the value of 1, which is the value of no smoking. So, there is insufficient evidence to conclude that the two groups are statistically significantly different.
Also, the p.values can confirm our found.


```{r}
#chisq test
chisq.test(new_bwgt$Smoking.f,new_bwgt$LowBirthWgt.f)
```

The chi-squared result shows that there is not enough evidence to say the low birth weight is causing by the smoking. 


```{r}
#oddsratio
oddsratio(new_bwgt$Smoking.f, new_bwgt$LowBirthWgt.f,rev=c("b"),method="wald")$measure
```

Even though the odd ratio of the non-smoking group is greater than one, the value is just 1.17. Also, the 95% C.I. does include 1, so our conclusion will be the same as above. We cannot say when women is smoking, it is more likely to deliver a low birth weight baby.


```{r}
# Fisher’s Exact Test 
tab22<- table(x=new_bwgt$Smoking.f , y=new_bwgt$LowBirthWgt.f)
tab22<- as.matrix(tab22)

# Two-sided alternative hypothesis
fisher.test(tab22)
# One-sided less than alternative hypothesis
fisher.test(tab22,alternative="less")
# One-sided greater than alternative hypothesis
fisher.test(tab22,alternative="greater")
```

All these three will lead to the same conclusion as we mentioned above, that is, there is no significant association at $\alpha=0.5$ between smoking and low birth weight. 


#
```{r}
# MH test:
mantelhaen.test(new_bwgt$Smoking.f, new_bwgt$LowBirthWgt.f, new_bwgt$Drinking.f,correct=T)
```

\vspace{1 in}
### Test for Association between Low Birth Weight and drinking
```{r}
#Risk ratio
RR1<-riskratio(x=new_bwgt$Drinking.f, y=new_bwgt$LowBirthWgt.f)
RR1$measure
RR1$p.value
```


```{r}
#chisq test
chisq.test(new_bwgt$Drinking.f,new_bwgt$LowBirthWgt.f)
```


```{r}
#oddsratio
oddsratio(new_bwgt$Drinking.f, new_bwgt$LowBirthWgt.f,rev=c("b"),method="wald")$measure
```


```{r}
# Fisher’s Exact Test 
tab23<- table(x=new_bwgt$Drinking.f, y=new_bwgt$LowBirthWgt.f)
tab22<- as.matrix(tab23)

# Two-sided alternative hypothesis
fisher.test(tab22)
# One-sided less than alternative hypothesis
fisher.test(tab22,alternative="less")
# One-sided greater than alternative hypothesis
fisher.test(tab22,alternative="greater")
```

```{r}
# MH test:
mantelhaen.test(new_bwgt$Drinking.f, new_bwgt$LowBirthWgt.f, new_bwgt$Smoking.f,correct=T)
```

All 5 tests' outputs give the same result, at $\alpha=0.5$,there is no strong evidence to conclude the relation between having a low birth weight baby and drinking is significant.


\newpage
# Code 3

### Test for Association between Low Birth Weight and Smoking controlling for Death
```{r}
new_bwgt1<-table(new_bwgt)
tab1<-margin.table(new_bwgt1, c(15,13,12))
d_yes<-tab1[, , Death.f = "No"]
d_no<-tab1[, , Death.f = "Yes"]
```


```{r}
#Risk ratio for Death= Yes
RR1<-riskratio(d_yes)
RR1$measure
RR1$p.value

#Risk ratio for Death= No
RR1<-riskratio(d_no)
RR1$measure
RR1$p.value

```


```{r}
#chisq test for Death= Yes
chisq.test(d_yes)

#chisq test for Death= No
chisq.test(d_no)

```


```{r}
#oddsratio for Death= Yes
oddsratio(d_yes,rev=c("b"),method="wald")$measure

#oddsratio for Death= No
oddsratio(d_no,rev=c("b"),method="wald")$measure
```


```{r}
# Fisher’s Exact Test for Death=Yes
# Two-sided alternative hypothesis
fisher.test(d_yes)
# One-sided less than alternative hypothesis
fisher.test(d_yes,alternative="less")
# One-sided greater than alternative hypothesis
fisher.test(d_yes,alternative="greater")


# Fisher’s Exact Test for Death=No
# Two-sided alternative hypothesis
fisher.test(d_no)
# One-sided less than alternative hypothesis
fisher.test(d_no,alternative="less")
# One-sided greater than alternative hypothesis
fisher.test(d_no,alternative="greater")
```

```{r}
# MH test:
mantelhaen.test(tab1,correct=T)
```


\vspace{0.5in}
### Test for Association between Low Birth Weight and Drinking controlling for Death
```{r}

tab2<-margin.table(new_bwgt1, c(15,11,12))
d_yes1<-tab2[, , Death.f = "No"]
d_no1<-tab2[, , Death.f = "Yes"]
```


```{r}
#Risk ratio for Death= Yes
RR1<-riskratio(d_yes1)
RR1$measure
RR1$p.value

#Risk ratio for Death= No
RR1<-riskratio(d_no1)
RR1$measure
RR1$p.value

```


```{r}
#chisq test for Death= Yes
chisq.test(d_yes1)

#chisq test for Death= No
chisq.test(d_no1)

```


```{r}
#oddsratio for Death= Yes
oddsratio(d_yes1,rev=c("b"),method="wald")$measure

#oddsratio for Death= No
oddsratio(d_no1,rev=c("b"),method="wald")$measure
```


```{r}
# Fisher’s Exact Test for Death=Yes
# Two-sided alternative hypothesis
fisher.test(d_yes1)
# One-sided less than alternative hypothesis
fisher.test(d_yes1,alternative="less")
# One-sided greater than alternative hypothesis
fisher.test(d_yes1,alternative="greater")


# Fisher’s Exact Test for Death=No
# Two-sided alternative hypothesis
fisher.test(d_no1)
# One-sided less than alternative hypothesis
fisher.test(d_no1,alternative="less")
# One-sided greater than alternative hypothesis
fisher.test(d_no1,alternative="greater")
```

```{r}
# MH test:
mantelhaen.test(tab2,correct=T)
```

According to the tests results shows above, we conclude the association between low birth weight and drinking controlling for death is not statistical significant.

\newpage
# Code 4

### Test for Association between Low Birth Weight and Death
```{r}
#Risk ratio
RR1<-riskratio(x=new_bwgt$Death.f, y=new_bwgt$LowBirthWgt.f)
RR1$measure
RR1$p.value
```
According to the risk ratio, since the 95% confidence interval does not contain 1, the association between the low birth weight and death is statistical significant.

```{r}
#chisq test
chisq.test(new_bwgt$Death.f,new_bwgt$LowBirthWgt.f)
```


```{r}
#oddsratio
oddsratio(new_bwgt$Death.f, new_bwgt$LowBirthWgt.f,rev=c("b"),method="wald")$measure
```


```{r}
# Fisher’s Exact Test 
tab24<- table(x=new_bwgt$Death.f , y=new_bwgt$LowBirthWgt.f)
tab22<- as.matrix(tab24)

# Two-sided alternative hypothesis
fisher.test(tab24)
# One-sided less than alternative hypothesis
fisher.test(tab24,alternative="less")
# One-sided greater than alternative hypothesis
fisher.test(tab24,alternative="greater")
```

```{r}
# MH test:
mantelhaen.test(new_bwgt$Death.f, new_bwgt$LowBirthWgt.f, new_bwgt$Drinking.f,correct=T)
```



